---
title: "Research seminar"
author: "SOFACK GHISLAIN(264513)"
date: "29/08/2020"
output: html_document
---
#Project title: Bioanalytical method Validation of Mass spectrometry analysis using R

#Supervisors: Frau Helga Weinscrott, Prof.Dr. Matthias Kohl



# Reading the data
```{r}
GA = read.csv2("GA.csv", header = TRUE, dec =",", stringsAsFactors= TRUE)  

# exracted just the data of the first analyte Glutaric Acid (GA) to work with.

```




# Creating a df for calculated conc. HSA 
```{r}

a = GA$Calculated.Concentration[grep("HSA", GA$Sample.Name)]

b = factor(GA$Sample.Name[grep("HSA", GA$Sample.Name)])



mat1 = do.call("rbind", split(a,b))

mat1[mat1 == 0] <- NA

View(mat1)


mat1 = data.frame(mat1)




mat1$Mean_HSA_uM = rowMeans(mat1, na.rm = TRUE) 

mat1$Stdabwn_uM = apply(mat1,1, sd, na.rm = TRUE) 

library("MKinfer")

CV <- function(x){ n <- length(x[!is.na(x)])
sqrt((n-1)/n)*sd( x, na.rm = TRUE)/mean(x, na.rm = TRUE)}

mat1$CV = apply((mat1[,1:(ncol(mat1)-2)]),1,CV)*100 




f <- function(x){ 
  n <- length(x[!is.na(x)])
  if(n>1) 
    res = ((sd(x, na.rm = TRUE) *qt(0.975, df = n-1))/(sqrt(n-1)*mean(x, na.rm = TRUE))) 
  else
    res = 0
  res
  } 

mat1$f = apply((mat1[,1:(ncol(mat1)-3)]),1,f)*100 


View(mat1)



```



# Creating a df for calculated S/N Human Serum Albumin 
```{r}


C = GA$Signal...Noise[grep("HSA", GA$Sample.Name)]

d = factor(GA$Sample.Name[grep("HSA", GA$Sample.Name)])

C = C[!d%in%c("HSA6", "HSA7")]

d = factor(d[!d%in%c("HSA6", "HSA7")])




mat2 = do.call("rbind", split(C,d))
mat2[mat2 == 0] <- NA

View(mat2)


mat2 = data.frame(mat2)


mat2$Mean_SN = rowMeans(mat2, na.rm = TRUE)

mat2$Stdabwn_uM = apply(mat2,1, sd, na.rm = TRUE) 

View(mat2)

```



# Creating a df for act vs calc conc S 
```{r}

f = c("Sd", "Sc", "Sb", "Sa", "S1", "S2", "S3", "S4", "S5", "S6", "S7")

ind = which(GA$Sample.Name%in%f) 
  
mat3 = data.frame(c(GA$Actual.Concentration[ind]),c(GA$Calculated.Concentration[ind]))
colnames(mat3) = c("act_conc_S", "calc_conc_S") 
rownames(mat3) = GA$Sample.Name[ind]

View(mat3)
```






# RECOVERY CALCULATION 

#1.) Linear adjustment of the calc.conc S vs act.conc S 
```{r}

f = c("Sa", "S1", "S2", "S3", "S4", "S5", "S6", "S7")

ind = which(rownames(mat3)%in%f) 


fit1 <- lm(mat3$calc_conc_S[ind] ~ mat3$act_conc_S[ind] )

summary(fit1) 


coef(fit1)[2] # Slope/steigung ~ 99% sensitivity (lowest concentration that can be measured with an acceptable limit of accuracy and precision)

summary(fit1)$coefficients[,4]  #p-values  sig. i.e the model is explained by the proposed regression

summary(fit1)$r.squared # R-squared (Recovery)  # r2>0.999 indicates recovery rate of close to 100% .

summary(fit1)$coefficients[2,2] # S.E   

(summary(fit1)$coefficients[2,2]/coef(fit1)[2])*100 ### S.E / slope (relativer Intra-Assay-Fehler der Wiederfindungsrate) # i.e <20% (FDA guideline)  --- <2% (accept)...the lower the r2 the grater the error and vice

  
# As expected, the recovery rates of analytes in acqeous solution is almost at 100 %



#Robust fit


library("robustbase")

fit1_rob = lmrob(mat3$calc_conc_S[ind] ~ mat3$act_conc_S[ind])
summary(fit1_rob)


coef(fit1_rob)[2] 

summary(fit1_rob)$coefficients[,4]  

summary(fit1_rob)$r.squared 

summary(fit1_rob)$coefficients[2,2]  

(summary(fit1_rob)$coefficients[2,2]/coef(fit1_rob)[2])*100 



```






#2.) Linear adjustment of the mean HSA vs act.conc S 
```{r}



f = c("Sd", "Sc", "Sb", "Sa", "S1", "S2", "S3", "S4", "S5", "S6", "S7")

ind = which(GA$Sample.Name%in%f) 


f1 = c("HSAd", "HSAc", "HSAb", "HSAa","HSA1", "HSA2", "HSA3", "HSA4", "HSA5", "HSA6", "HSA7")

ind1 = which(rownames(mat1)%in%f1) 


mat4 = data.frame(c(GA$Actual.Concentration[ind]),c(GA$Calculated.Concentration[ind]), c(mat1$Mean_HSA_uM[ind1]))

colnames(mat4) = c("act_conc_S", "calc_conc_S", "mean_HSA_uM")


View(mat4)



fit2 <- lm(mat4$mean_HSA_uM ~ mat4$act_conc_S)

summary(fit2)

coef(fit2)[2] # steigung/coefficients ~ 98% sensistivity

summary(fit2)$coefficients[,4]  #pvals

summary(fit2)$r.squared # R-squared ~ strong positiv


summary(fit2)$coefficients[2,2] # S.E

(summary(fit2)$coefficients[2,2]/coef(fit2)[2])*100 # S.E / slope  ~ < 1% (very accurate)

# Recovery rate of GA in serum/ plasma is found to be at 99% (good). Systematic deviations, such as with citric acid,  indicate that there are effects in the real samples that  prevent the analyte from being extracted 100%:  e.g. it may be that citric acid binds to various proteins etc. in the biological plasma and therefore cannot be extracted as well.

# FDA gudeline: The recovery need not to be 100%, but should be concise, precise and reproducible.








#Robust fit


fit2_rob = lmrob(mat4$mean_HSA_uM ~ mat4$act_conc_S)
summary(fit2_rob)


coef(fit2_rob)[2] 

summary(fit2_rob)$coefficients[,4]  

summary(fit2_rob)$r.squared 

summary(fit2_rob)$coefficients[2,2]  

(summary(fit2_rob)$coefficients[2,2]/coef(fit2_rob)[2])*100 



```






# Ploting the Recovery (calc.conc.S and calc.conc.HSA against act_conc_ S )
```{r}

x = mat3$act_conc_S[ind]
y1=mat4$mean_HSA_uM[ind1]
y2=mat3$calc_conc_S[ind]

plot(x,y1)
plot(x,y2)

data1 = cbind(x,y1)
data2 = cbind(x, y2)


plot(y2 ~ x, data=data2,col='blue')
abline(lm(y2 ~ x),col ='blue')

points(y1 ~ x, data = data1,col='orange')
abline(lm(y1 ~ x),col='orange')

# linear range is from 0-50 uM
```






#3.) Linear adjustment of the mean HSA vs calc.conc S
```{r}

f = c("Sa", "S1", "S2", "S3", "S4", "S5", "S6", "S7")

ind = which(GA$Sample.Name%in%f) 


f1 = c("HSAa","HSA1", "HSA2", "HSA3", "HSA4", "HSA5", "HSA6", "HSA7")

ind1 = which(rownames(mat1)%in%f1) 



fit3 <- lm(mat4$mean_HSA_uM[ind1] ~ mat4$calc_conc_S[ind])

summary(fit3)

coef(fit3)[2] # steigung / slope ~ 99% sensitive

summary(fit3)$coefficients[,4]  #pvals

summary(fit3)$r.squared # R-squared ...Internal criteria is that r2 !< 0.99

summary(fit3)$coefficients[2,2] # S.E

(summary(fit3)$coefficients[2,2]/coef(fit3)[2])*100 # S.E / slope = >2%


# The analyte GA can be extracted with acceptable limit of accuracy and precision





#Robust fit


fit3_rob = lmrob(mat4$mean_HSA_uM[ind1] ~ mat4$calc_conc_S[ind])

summary(fit3_rob)

coef(fit3_rob)[2] 

summary(fit3_rob)$coefficients[,4]  

summary(fit3_rob)$r.squared 

summary(fit3_rob)$coefficients[2,2]  

(summary(fit3_rob)$coefficients[2,2]/coef(fit3_rob)[2])*100 






```







# Calculation of LOQ (CV = 15% and 20%) 

## y = ax^b
## log(y) = log(a) + blog(x)
## intercept = log(a)
## slope = b


```{r}



f = c("HSA1", "HSA2", "HSA3", "HSA4", "HSA5", "HSA6", "HSA7")

ind = which(rownames(mat1)%in%f) 




fit5 <- lm(log(mat1$CV[ind]) ~ log(mat1$Mean_HSA_uM[ind]))

summary(fit5)

# from the data CV are high at lower conc of HSA and vice versa. So here, we could probably exclude outliers using the robust fits


#Robust fit


fit5_rob = lmrob(log(mat1$CV[ind]) ~ log(mat1$Mean_HSA_uM[ind]))

summary(fit5_rob)

coef(fit5_rob)[2] 

summary(fit5_rob)$coefficients[,4]  

summary(fit5_rob)$r.squared 

summary(fit5_rob)$coefficients[2,2]  

(summary(fit5_rob)$coefficients[2,2]/coef(fit5_rob)[2])*100 








loq15 = function(x){
  exp((log(0.15/(exp(coef(x)[1])/100))/coef(x)[2]))
}
loq15(fit5) # conc. that can be determined at CV=15% 


loq20 = function(x){
  exp((log(0.20/(exp(coef(x)[1])/100))/coef(x)[2]))
}
loq20(fit5) # conc. that can be determined at CV = 20%


library(ggplot2)
ggplot(mat1) + 
  geom_point(aes(y = CV, x = Mean_HSA_uM)) 

```






# Calculation of the LOD and LOQ (S/N ratio)

##1.) Linear fit (lineare Anpassung)
```{r}

f = c("HSA0","HSAd", "HSAc", "HSAb", "HSAa", "HSA1", "HSA2", "HSA3", "HSA4")

ind = which(rownames(mat1)%in%f) 



f1 = c("HSA0","HSAd", "HSAc", "HSAb", "HSAa", "HSA1", "HSA2", "HSA3", "HSA4")

ind1 = which(rownames(mat2)%in%f1) 



mat5 = data.frame(mat2$Mean_SN[ind1], mat1$Mean_HSA_uM[ind])

colnames(mat5) = c("Mean_SN", "Mean_HSA_uM")

fit6 = lm(mat5$Mean_SN ~mat5$Mean_HSA_uM) # Only the concentrations up to the 4th standard above the zero sample (HSA0) were taken into account because the increase in S / N values is proportional to the concentrations only in this range

summary(fit6)

coef(fit6)[1] # Intercept

coef(fit6)[2] # Slope/steigung 





#Robust fit


fit6_rob = lmrob(mat5$Mean_SN ~ mat5$Mean_HSA_uM)

summary(fit6_rob)

coef(fit6_rob)[2] 

summary(fit6_rob)$coefficients[,4]  

summary(fit6_rob)$r.squared 

summary(fit6_rob)$coefficients[2,2]  

(summary(fit6_rob)$coefficients[2,2]/coef(fit6_rob)[2])*100 







lod = function(x){
  (3-coef(x)[1])/coef(x)[2]
}

lod(fit6) # lowest quantity of analyte that can be distinguished from the noise i.e 3 sigma away from blank


lloq = function(x){
  (10-coef(x)[1])/coef(x)[2]
}
lloq(fit6) # conc. which gives a signal as high as 10x noise of the blank


```





##2.) Potenz Anpassung (exponential model) 
# y = (Bo*B1^lnX) * E
# log(y) = logBo + logB1*logX + log E
```{r}




fit7 = lm(log(mat5$Mean_SN) ~ I(log(mat5$Mean_HSA_uM)))

summary(fit7)

#logBo = 2.51306
# Bo = exp(2.51306)





#Robust fit


fit7_rob = lmrob(log(mat5$Mean_SN) ~ I(log(mat5$Mean_HSA_uM)))

summary(fit7_rob)

coef(fit7_rob)[2] 

summary(fit7_rob)$coefficients[,4]  

summary(fit7_rob)$r.squared 

summary(fit7_rob)$coefficients[2,2]  

(summary(fit7_rob)$coefficients[2,2]/coef(fit7_rob)[2])*100 






lod_pa = function(x){
  exp((log(3/exp(coef(x)[1]))/coef(x)[2]))
}

lod_pa(fit7)

lloq_pa = function(x){
  exp((log(10/exp(coef(x)[1]))/coef(x)[2]))
}

lloq_pa(fit7)


# Linear fit is better than the exponential fit

```







##3.) Quadratische Anpassung (polynomial regression model)
# y = β0 + β1*X + β2*X^2 + E
```{r}

fit8 = lm(mat5$Mean_SN ~ mat5$Mean_HSA_uM + I(mat5$Mean_HSA_uM^2))

summary(fit8)




#Robust fit


fit8_rob = lmrob(mat5$Mean_SN ~ mat5$Mean_HSA_uM + I(mat5$Mean_HSA_uM^2))

summary(fit8_rob)

coef(fit8_rob)[2] 

summary(fit8_rob)$coefficients[,4]  

summary(fit8_rob)$r.squared 

summary(fit8_rob)$coefficients[2,2]  

(summary(fit8_rob)$coefficients[2,2]/coef(fit8_rob)[2])*100 





lod_qa = function(x){
 n = -3 + coef(x)[1]
  (-coef(x)[2] + sqrt((coef(x)[2]^2) - 4*coef(x)[3]*n))/ (2*coef(x)[3])
}

lod_qa(fit8)


lloq_qa = function(x){
  m = -10 + coef(x)[1]
  (-coef(x)[2] + sqrt((coef(x)[2]^2) - 4*coef(x)[3]*m))/ (2*coef(x)[3])
}

lloq_qa(fit8)



```






# Plotting S/N HAS
```{r}
##############################

ggplot(mat5,aes(mat1$Mean_HSA_uM[1:9], mat2$Mean_SN[1:9])) + labs(title = paste("S/N HAS"), x = "mean Human Serum Albumin uM", y = "S/N Human Serum Albumin") +
  geom_point() + stat_smooth(method = "lm",
    col = "black",
    se = FALSE,
    size = .5)
##############################

```








# Intra-Assay precision for calculaing recovery (resumee sheet excel.)
```{r}


iap <- data.frame(c(summary(fit1)$r.squared, 
  summary(fit1)$coefficients[2,2], summary(fit1)$r.square, 
  (summary(fit1)$coefficients[2,2]/coef(fit1)[2])*100,
  (summary(fit1)$r.squared)*100,  loq15(fit5), lloq(fit6), lloq(fit8),
  0, (lloq(fit6) + lloq(fit8) + 0)/2 , 0))

rownames(iap) = c("Wiederfindungsrate","Fehler.abs","Regrssionskoeffizi
  ent","%F Intra-Assay","Wiederfindungsrate%", "LOQ(CV15%)", 
  "LOQ(S/N=10)", "LOQ_Quadfit", "LOQ(S/N=10)potfit","LOQmean(S/N=10)", "LOD(S/N = 3)PotFit")

colnames(iap) =c("Glutaric Acid")

iap
```

